---
title: "FFM Comparison"
author: "R. Peek"
date: "Updated `r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  bookdown::html_document2:
    theme: cosmo
    number_sections: false
    highlight: textmate
    toc: true
    toc_float:
        collapsed: true
    code_fold: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("R/packages.R")
source("R/data_plot_summarize.R")

# load the data
basedir <- glue("{here()}/output")
df_all <- fst::read_fst(glue("{basedir}/ffm_combined_tidy.fst"))
```

To better understand how metrics have changed between previous iterations of the Functional Flow Calculator, here are some comparison plots across stream class and metric. 

<br>

## Summarizing Old Vs New 

1. I pulled all the stream class data in from the raw stream `Class` folders using this code to read, tidy, and clean all the data and get it into one single file:

 - [`get_data.R`](https://raw.githubusercontent.com/ryanpeek/ffm_comparison/main/R/get_data.R)
 
2. I then combined each "version" (2019 and 2020) of the data using this code:

 - [`combine_data.R`](https://raw.githubusercontent.com/ryanpeek/ffm_comparison/main/R/combine_data.R)
 
3. Once we have a single file with all the data, we can evaluate, plot, and summarize as we like.

 - [`plot_summarize.R`](https://raw.githubusercontent.com/ryanpeek/ffm_comparison/main/R/plot_summarize.R)
 
### Basic Workflow
 
```{r workflow, eval=F, echo=T}
# PACKAGES
source("R/packages.R")

# FUNCTION TO GET DATA FROM RAW SOURCE AND TIDY
source("R/get_data.R")

# use returndata=TRUE to return raw tidied data to global environment
get_data(2019, returndata = FALSE)
get_data(2020, returndata = FALSE)

# FUNCTION TO COMBINE VERSIONS (2019 and 2020) INTO SINGLE DATAFRAME
source("R/combine_data.R") # automatically loads data

# PLOT/SUMMARIZE
source("R/plot_summarize.R")

# get table for stream class and flow Chx
# flow chx: "Timing" "Magnitude" "Duration" "Rate of change" "Frequency"
flow_chx_dt(flow_characteristic="Magnitude") # pass a stream class to the function
plot_flow_chx(df_all, stream_class = "Class-1", flow_characteristic="Magnitude")

```


There are 9 stream classes, and 5 stream characteristics (**Timing, Magnitude, Duration, Rate of Change and Frequency**).

<br>

## Magnitude

If we look at Magnitude metrics for all Stream classes, and compare each side by side, we can quickly see which metrics are identical and which have changed. This is the **mean value** for FFM by stream class and flow metric across all gages within that stream class.

Changed: 

 - **Peak_2**
 - **Peak_5**

```{r, echo=F, eval=T, message=FALSE, warning=FALSE}

flow_chx_dt(data = df_all, flow_characteristic = "Magnitude")

```

### Stream Class-1

If we look at Stream Class-1 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(data = df_all, "Class-1", "Magnitude")

```

### Stream Class-2

If we look at Stream Class-2 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-2", "Magnitude")

```

### Stream Class-3

If we look at Stream Class-3 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-3", "Magnitude")

```

### Stream Class-4

If we look at Stream Class-4 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-4", "Magnitude")

```

### Stream Class-5

If we look at Stream Class-5 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-5", "Magnitude")

```

### Stream Class-6

If we look at Stream Class-6 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-6", "Magnitude")

```

### Stream Class-7

If we look at Stream Class-7 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-7", "Magnitude")

```

### Stream Class-8

If we look at Stream Class-8 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-8", "Magnitude")

```

### Stream Class-9

If we look at Stream Class-9 for Magnitude and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-9", "Magnitude")

```




<br>

## Timing

If we look at Timing metrics for all Stream classes, and compare each side by side, we can quickly see there are **NO DIFFERENCES**.

```{r, echo=F, eval=T, message=FALSE, warning=FALSE}

flow_chx_dt(data = df_all, flow_characteristic = "Timing")

```

<br>

## Duration

If we look at Duration metrics for all Stream classes, and compare each side by side, we can quickly see which metrics are identical and which have changed. This is the **mean value** for FFM by stream class and flow metric across all gages within that stream class.

Changed: 

 - **Peak_Dur_2**
 - **Peak_Dur_5**

```{r, echo=F, eval=T, message=FALSE, warning=FALSE}

flow_chx_dt(data=df_all, flow_characteristic = "Duration")

```

### Stream Class-1

If we look at Stream Class-1 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-1", "Duration")

```

### Stream Class-2

If we look at Stream Class-2 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-2", "Duration")

```

### Stream Class-3

If we look at Stream Class-3 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-3", "Duration")

```

### Stream Class-4

If we look at Stream Class-4 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-4", "Duration")

```

### Stream Class-5

If we look at Stream Class-5 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-5", "Duration")

```

### Stream Class-6

If we look at Stream Class-6 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-6", "Duration")

```

### Stream Class-7

If we look at Stream Class-7 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-7", "Duration")

```

### Stream Class-8

If we look at Stream Class-8 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-8", "Duration")

```

### Stream Class-9

If we look at Stream Class-9 for Duration and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-9", "Duration")

```




<br>

## Rate of Change

If we look at Rate of change metrics for all Stream classes, and compare each side by side, we can quickly see there are **NO DIFFERENCES**.

```{r, echo=F, eval=T, message=FALSE, warning=FALSE}

flow_chx_dt(df_all, flow_characteristic = "Rate of change")

```

<br>

## Frequency

If we look at Frequency metrics for all Stream classes, and compare each side by side, we can quickly see which metrics are identical and which have changed. This is the **mean value** for FFM by stream class and flow metric across all gages within that stream class.

Changed: 

 - **Peak_Fre_2**
 - **Peak_Fre_5**

```{r, echo=F, eval=T, message=FALSE, warning=FALSE}

flow_chx_dt(df_all, flow_characteristic = "Frequency")

```

### Stream Class-1

If we look at Stream Class-1 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-1", "Frequency")

```

### Stream Class-2

If we look at Stream Class-2 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-2", "Frequency")

```

### Stream Class-3

If we look at Stream Class-3 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-3", "Frequency")

```

### Stream Class-4

If we look at Stream Class-4 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-4", "Frequency")

```

### Stream Class-5

If we look at Stream Class-5 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-5", "Frequency")

```

### Stream Class-6

If we look at Stream Class-6 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-6", "Frequency")

```

### Stream Class-7

If we look at Stream Class-7 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-7", "Frequency")

```

### Stream Class-8

If we look at Stream Class-8 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-8", "Frequency")

```

### Stream Class-9

If we look at Stream Class-9 for Frequency and visualize these data (so combine across gages within a stream class):

```{r, echo=F, eval=T, fig.retina=2, message=FALSE, warning=FALSE}

plot_flow_chx(df_all, "Class-9", "Frequency")

```



## Resources & Websites {.tabset .tabset-fade .tabset-pills}

### Python FFC

The original python repository where the current up-to-date version of the FF calculator:

 - [FFC On Github](https://github.com/leogoesger/func-flow)
 
### Eflows Website

The [eflows.ucdavis.edu](https://eflows.ucdavis.edu) website uses the python code to generate functional flow metrics. This is the visual "GUI" that many folks are using currently.

 - [eflows.ucdavis.edu](https://eflows.ucdavis.edu)

### R API Wrapper

This is an R package that essentially uses the underlying code from the **eflows website** to calculate FFM in R.

 - [R ffc package](https://github.com/ceff-tech/ffc_api_client)

<br>
